---
title: "`countries`"
tags: [r]
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

`countries` offers a consistent framework for working with country names and 
codes, with functions for parsing irregular country names and converting codes 
and names in many languages and formats.


## Installation

Install from GitHub with

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("alistaire47/countries")
```

---

# A painless country name and code workflow

Working with country data can be frustrating. Even with well-curated data like 
[`gapminder`](https://github.com/jennybc/gapminder), there are some oddities:

```{r, message=FALSE}
library(countries)
library(gapminder)
library(tidyverse)    # Works equally well in any grammar.

grep('Korea', unique(gapminder$country), value = TRUE)
grep('Yemen', unique(gapminder$country), value = TRUE)
```

`countries` offers a framework for working with country names and codes without 
manually editing data or scraping codes from Wikipedia.


## 1. Standardize

If data has non-standardized names, standardize them to an ISO 3166-1 code 
or other standardized code or name with `parse_country`:

```{r}
gap <- gapminder %>% 
    # standardize to ISO 3166 Alpha-2 code
    mutate(country_code = parse_country(country))

gap %>%
    select(country, country_code, year, lifeExp) %>%
    sample_n(10)
```


## 2. Convert

If data comes with countries already coded, convert them with 
`as_country_code()`:

```{r, message = FALSE}
# 2016 Olympic gold medal data
olympics <- read_tsv('https://raw.githubusercontent.com/nbremer/olympicfeathers/gh-pages/data/raw%20medal%20data/Rio%202016%20gold%20medal%20winners.txt')

olympics %>% count(country = as_country_code(NOC, from = 'ioc'), sort = TRUE)
```

or to convert to country names, use `as_country_name()`:

```{r}
olympics %>% 
    count(country = as_country_name(NOC, from = 'ioc'), 
          Event_gender) %>% 
    spread(Event_gender, n) %>% 
    arrange(desc(W))
```

or translate to another language:

```{r}
olympics$NOC %>% unique() %>% 
    as_country_name(from = 'ioc', to = 'ja') %>% 
    head(10)
```


## 3. Format

A particularly common hangup with country data is presentation. While 
"Yemen, Rep." may be fine for exploratory work, to create a plot to share, 
such names need to be changed to something more palatable either by editing 
the data or manually overriding the labels directly on the plot.

If the existing format is already standardized, `countries` offers another 
option: use a formatter function created with `format_country`, just like for 
thousands separators or currency formatting:

```{r}
gap %>% 
    group_by(country_code) %>% 
    summarise(start_life_exp = lifeExp[which.min(year)], 
              stop_life_exp = lifeExp[which.max(year)], 
              diff_life_exp = stop_life_exp - start_life_exp) %>% 
    top_n(10, diff_life_exp) %>%
               # Plot country codes...
    ggplot(aes(x = country_code,
               ymin = start_life_exp, 
               ymax = stop_life_exp, 
               colour = country_code)) + 
    geom_linerange(size = 20, show.legend = FALSE) + 
                     # ...just pass `labels` a formatter function!
    scale_x_discrete(labels = format_country()) + 
    labs(title = 'Life gets better',
         subtitle = 'Largest increase in life expectancy',
         x = NULL, y = 'Life expectancy')
```

---


## Data

The data underlying `countries` comes from a number of sources, including

- [The Unicode Common Locale Data Repository (CLDR)
Project](http://cldr.unicode.org/) supplies country names in many, many
languages, from Afrikaans to Zulu. Even better, [CLRDR aspires to use the most
customary name](http://cldr.unicode.org/translation/country-names) instead of
formal or official ones, e.g. "Switzerland" instead of "Swiss Confederation". 
See included LICENSE for terms of use.
- [The CIA World 
Factbook](https://www.cia.gov/library/publications/the-world-factbook/index.html)
supplies a standardized set of names and codes. 
- [The National Geospatial-Intelligence Agency 
(NGA)](http://geonames.nga.mil/gns/html/countrycodes.html) is the organization 
responsible for standardizing US government use of country codes. It inherited 
the now-deprecated FIPS 10-4 from NIST, which it turned into the GEC, which is 
now also deprecated in favor of GENC, a US government profile of ISO 3166.
- Yes, [Wikipedia](https://en.wikipedia.org/wiki/Category:Lists_of_country_codes) 
offers a rich set of country codes, if aggregated into a useful format.
- [Open Knowledge International's Frictionless Data](http://data.okfn.org/data/core/country-codes) 
supplies a set of codes collated from a number of sources.
- The regex powering `parse_country()` are from
[`countrycode`](https://github.com/vincentarelbundock/countrycode). If you would
like to improve both packages, please contribute there!


## License

`countries` is licenced as open-source software under
[GPL-3](https://www.gnu.org/licenses/gpl.html).
