---
title: "passport"
subtitle: "Travel smoothly between country name and code formats"
author: "Edward Visel"
date: "`r format(Sys.Date(), '%d %B %Y')`"
pkgdown:
  as_is: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{passport}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- passport.md is generated from passport.Rmd. Please edit that file -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`passport` smooths the process of working with country names and codes via
powerful parsing, standardization, and conversion utilities arranged in a
simple, consistent API. Country name formats include multiple sources including
the Unicode CLDR common-sense standardizations in hundreds of languages.

Working with country data can be frustrating. Even with well-curated data like 
[`gapminder`](https://github.com/jennybc/gapminder), there are some oddities:

```{r intro, message=FALSE}
library(passport)
library(gapminder)
library(dplyr)    # Works equally well in any grammar.
library(tidyr)
set.seed(47)

grep("Korea", unique(gapminder$country), value = TRUE)
grep("Yemen", unique(gapminder$country), value = TRUE)
```

`passport` offers a framework for working with country names and codes without 
manually editing data or scraping codes from Wikipedia.


## I. Standardize

If data has non-standardized names, standardize them to an ISO 3166-1 code 
or other standardized code or name with `parse_country`:

```{r standardize-1}
gap <- gapminder %>% 
    # standardize to ISO 3166 Alpha-2 code
    mutate(country_code = parse_country(country))

gap %>%
    select(country, country_code, year, lifeExp) %>%
    sample_n(10)
```

If country names are particularly irregular, in unsupported languages, or are 
even just unique location names, `parse_country` can use Google Maps or Data 
Science Toolkit geocoding APIs to parse instead of regex:

```{r standardize-2, eval=FALSE}
parse_country(c("somewhere in Japan", "日本", "Japon", "जापान"), how = "google")
#> [1] "JP" "JP" "JP" "JP"

parse_country(c("1600 Pennsylvania Ave, DC", "Eiffel Tower"), how = "google")
#> [1] "US" "FR"
```


### II. Convert

If data comes with countries already coded, 

- convert them to ISO or other codes with `as_country_code()`
- convert them to country names with `as_country_name()`
- convert them to other languages with `as_country_name()`

```{r convert-1, message = FALSE}
# NATO member defense expenditure data; see `?nato`
data("nato", package = "passport")

nato %>% 
    select(country_stanag) %>% 
    distinct() %>%
    mutate(
        country_iso = as_country_code(country_stanag, from = "stanag"),
        country_name = as_country_name(country_stanag, from = "stanag", short = FALSE),
        country_name_thai = as_country_name(country_stanag, from = "stanag", to = "ta-my")
    )
```

Language formats largely follow [IETF language tag BCP
47](https://en.wikipedia.org/wiki/IETF_language_tag) format. For all available
formats, run `DT::datatable(codes)` for an interactive widget of format names
and further information.

## III. Format

A particularly common hangup with country data is presentation. While 
"Yemen, Rep." may be fine for exploratory work, to create a plot to share, 
such names need to be changed to something more palatable either by editing 
the data or manually overriding the labels directly on the plot.

If the existing format is already standardized, `passport` offers another 
option: use a formatter function created with `country_format`, just like for 
thousands separators or currency formatting. Reorder simply with 
`order_countries`:

```{r format, fig.width=5}
library(ggplot2)

living_longer <- gap %>% 
    group_by(country_code) %>% 
    summarise(start_life_exp = lifeExp[which.min(year)], 
              stop_life_exp = lifeExp[which.max(year)], 
              diff_life_exp = stop_life_exp - start_life_exp) %>% 
    top_n(10, diff_life_exp) 

# Plot country codes...
ggplot(living_longer, aes(x = country_code, y = stop_life_exp - 4.5,
                          ymin = start_life_exp, 
                          ymax = stop_life_exp - 4.5, 
                          colour = factor(diff_life_exp))) + 
    geom_point(pch = 17, size = 7) + 
    geom_linerange(size = 5) + 
                     # ...just pass `labels` a formatter function!
    scale_x_discrete(labels = country_format(),
                     # Easily change order
                     limits = order_countries(living_longer$country_code, 
                                              living_longer$diff_life_exp)) + 
    scale_y_continuous(limits = c(30, 80)) + 
    labs(title = "Life gets better",
         subtitle = "Largest increase in life expectancy",
         x = NULL, y = "Life expectancy") + 
    theme(axis.text.x = element_text(angle = 30, hjust = 1), 
          legend.position = "none")
```

By default `country_format` will use Unicode CLDR (see below) English names, 
which are intelligible and suitable for most purposes. If desired, other 
languages or formats can be specified just like in `as_country_name`.

---


## Data

The data underlying `passport` comes from a number of sources, including

- [The Unicode Common Locale Data Repository (CLDR)
Project](http://cldr.unicode.org/) supplies country names in many, many
languages, from Afrikaans to Zulu. Even better, [CLDR aspires to use the most
customary name](http://cldr.unicode.org/translation/displaynames/country-names) instead of
formal or official ones, e.g. "Switzerland" instead of "Swiss Confederation".
- [The United Nations Statistics 
Division](https://unstats.un.org/unsd/methodology/m49/overview/) maintains and 
publishes the M.49 region code and the UN geoscheme region codes and names.
- [The CIA World 
Factbook](https://www.cia.gov/library/publications/the-world-factbook/index.html)
supplies a standardized set of names and codes. 
- [The National Geospatial-Intelligence Agency 
(NGA)](http://geonames.nga.mil/gns/html/countrycodes.html) is the organization 
responsible for standardizing US government use of country codes. It inherited 
the now-deprecated FIPS 10-4 from NIST, which it turned into the GEC, which is 
now also deprecated in favor of GENC, a US government profile of ISO 3166.
- [Wikipedia](https://en.wikipedia.org/wiki/Category:Lists_of_country_codes) 
offers a rich set of country codes, some of which are aggregated here.
- Open Knowledge International's Frictionless Data supplies [a set of codes 
collated from a number of sources](https://www.datahub.io/core/country-codes) on 
datahub.io.
- The regex powering `parse_country()` are from
[`countrycode`](https://github.com/vincentarelbundock/countrycode). If you 
would like to improve both packages, please contribute regex there!


## Licensing

`passport` is licensed as open-source software under
[GPL-3](https://www.gnu.org/licenses/gpl.html). Unicode CLDR data is licensed 
according to [its own 
license](https://github.com/unicode-cldr/cldr-json/blob/master/LICENSE), a copy 
of which is included. `countrycode` regex are used as a modification under 
GPL-3; see the included aggregation script for modifiying code and date.
